Classification of shapes

### 2019-06-24 manuscript/Manuscript2.Rnw
Manuscript.

@@

\subsection{Notation}
A closed planar curve $C$ is always an absolutely continuous mapping $C:\uc \to \Rtwo$, and the set of planar closed curves will be denoted by $\mathcal{C}$. The set $\mathcal{C}$ is equipped with the norm $\|x\|_{\ltwo}:=[\int_\uc \|x(t)\|_2dt]^{1/2}$ where $\|\cdot\|_2$ is the Euclidean norm in $\Rtwo$. $SO(2)$ is the special orthogonal group of rotation matrices of $\Rtwo$, and $\Gamma$ is group of orientation-preserving diffeomorphisms of $\uc$. The set $\Gamma_I$ will denote the group $\{\gamma:[0,1]\to[0,1], \gamma'>0,\gamma(0)=0, \gamma(1)=1\}$. 
\subsection{Square-root velocity transform}
For a detailed introduction to the transform, its properties and advantages we refer to the reader to Chapter 6 of the book by \cite{AK}. Here we briefly outline the important concepts required for our purposes. 
For an absolutely continuous curve $\C:\uc \to \Rtwo$ consider the transformation 
$$\C \mapsto \frac{C'}{\|C'\|_{\ltwo}}=:Q_\C,$$ 
where $\C'$ is the (vector) derivative of $\C(t)$ with respect to $t$, and $\|\cdot\|$ is the usual Euclidean norm in $\Rtwo$.

@@

Two important ramifications of the described framework, motivating its use in our work for analyzing shapes of curves, are the following. Under the square-root velocity framework:
\begin{enumerate}[1.]
\item The actions of $SO(2)$ and $\Gamma$ on $\mathcal{Q}$ commute, i.e. they can be applied to a curve in any order. This ensures that their combined action is given by the product group $\Gamma \times SO(2)$.
\item The action of $\Gamma \times SO(2)$ on $\mathcal{Q}$ is by isometries: Given two curves $\C_1$ and $\C_2$, we have $\|OQ_{\C_1}(\gamma)\sqrt{\gamma '}-OQ_{\C_2}(\gamma)\sqrt{\gamma '}\|_{\ltwo}=\|Q_{C_1}-Q_{C_2}\|_{\ltwo}$, for every $(\gamma,O) \in \Gamma \times SO(2)$. This ensures that if two square-root velocity transformed curves are rotated and reparameterized the same way, their distance remains unchanged. 
\end{enumerate}
 Starting with a curve $\C$ we can now define its shape to be the equivalence class or its orbit of its corresponding square-root transform:
 $$[Q_\C]=\text{closure}\{OQ_\C(\gamma) \sqrt{\gamma'}: (\gamma,O) \in \Gamma \times SO(2)\},$$
 where the closure is with respect to the norm $\|\cdot\|_{\ltwo}$ on $Q$. The shape space consequently is defined as $\mathcal{Q}_s:=\{[Q_\C]: Q_\C \in \mathcal{Q}\}$. Property (2) ensures that the metric induced by the norm $\|\cdot\|_{\ltwo}$ on $\mathcal{Q}$ descends onto a metric $d$ on the shape space (quotient space) $\mathcal{Q}_s$ in a natural way. Given two curves $\C_1$ and $\C_2$, the shape distance between them is defined as
 \begin{align}
 \label{distance}
 d(\C_1,\C_2)&:=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|Q_{\C_1} - OQ_{\C_2}(\gamma)\sqrt{\gamma'}\|_{\ltwo}\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)}\Big[\int_{\uc}\left\|Q_{C_1}(t)-OQ_{C_2}(\gamma(t))\sqrt{\gamma'(t)}\right\|_2^2dt\Big]^{1/2} \nonumber\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|OQ_{C_1}(\gamma)\sqrt{\gamma'} - Q_{C_2}\|_{\ltwo}\nonumber.
  \end{align}
  The symmetry with respect to the action either on $Q_{C_1}$ or $Q_{C_2}$ is an attractive feature and will be used profitably in the sequel. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Curve completion and classification}
Suppose we are given closed planar curves $\C^m_j:\uc \to \mathbb{R}^2,j=1\ldots,n$ each of which has been observed only on a region $\R_j \subset \uc$, where $\uc$ is the unit circle in $\Rtwo$. We assume that the curves are absolutely continuous. Additionally, a training sample $\{(y_i,\C^o_i),1,\ldots,N\}$ consisting of class labels $y_i \in \{0,1\}$ and $C^o_i:\uc\to \mathbb{R}^2$, fully observed on a common domain $\uc$, is provided. The set of fully observed curves are elements of the set $\mathcal{C}$; denote by $\mathcal{C}^m$ the set of all partially observed curves.

>>>>

In this instance, it is also of interest to preserve the size and shape of the curve.  This simply requires removing any scale changes from the definition of the equivalence class.  This size-and-shape of a parameterized $C: D \rightarrow \mathbb{R}^2$ is defined to be the equivalence class:

$$
[C^{\star}] :=  \{\sigma O C(\gamma(t)) + a, \gamma \in \Gamma, O \in SO(2), a\in \mathbb{R}^2 \} 
$$

Thus $[C^{\star}]$ is the set of all possible curves that can be obtained through a translation ($C+a$), rotation ($OC$), a reparameterization ($C(\gamma)$), or any combination of the transofrmations of curve $C$, notably with scaling not included as one of the possible transformations.

@@

\subsection{Notation}
A closed planar curve $C$ is always an absolutely continuous mapping $C:\uc \to \Rtwo$, and the set of planar closed curves will be denoted by $\mathcal{C}$. The set $\mathcal{C}$ is equipped with the norm $\|x\|_{\ltwo}:=[\int_\uc \|x(t)\|_2dt]^{1/2}$ where $\|\cdot\|_2$ is the Euclidean norm in $\Rtwo$. $SO(2)$ is the special orthogonal group of rotation matrices of $\Rtwo$, and $\Gamma$ is group of orientation-preserving diffeomorphisms of $\uc$. The set $\Gamma_I$ will denote the group $\{\gamma:[0,1]\to[0,1], \gamma'>0,\gamma(0)=0, \gamma(1)=1\}$.

\subsection{Square-root velocity transform}
For a detailed introduction to the transform, its properties and advantages we refer to the reader to Chapter 6 of the book by \cite{AK}. Here we briefly outline the important concepts required for our purposes.

For an absolutely continuous curve $\C:\uc \to \Rtwo$ consider the transformation 
$$\C \mapsto \frac{C'}{\|C'\|_{\ltwo}}=:Q_\C,$$ 
where $\C'$ is the (vector) derivative of $\C(t)$ with respect to $t$, and $\|\cdot\|$ is the usual Euclidean norm in $\Rtwo$.

@@

Two important ramifications of the described framework, motivating its use in our work for analyzing shapes of curves, are the following. Under the square-root velocity framework:
\begin{enumerate}[1.]
\item The actions of $SO(2)$ and $\Gamma$ on $\mathcal{Q}$ commute, i.e. they can be applied to a curve in any order. This ensures that their combined action is given by the product group $\Gamma \times SO(2)$.
\item The action of $\Gamma \times SO(2)$ on $\mathcal{Q}$ is by isometries: Given two curves $\C_1$ and $\C_2$, we have $\|OQ_{\C_1}(\gamma)\sqrt{\gamma '}-OQ_{\C_2}(\gamma)\sqrt{\gamma '}\|_{\ltwo}=\|Q_{C_1}-Q_{C_2}\|_{\ltwo}$, for every $(\gamma,O) \in \Gamma \times SO(2)$. This ensures that if two square-root velocity transformed curves are rotated and reparameterized the same way, their distance remains unchanged. 
\end{enumerate}
 Starting with a curve $\C$ we can now define its shape to be the equivalence class or its orbit of its corresponding square-root transform:
 $$[Q_\C]=\text{closure}\{OQ_\C(\gamma) \sqrt{\gamma'}: (\gamma,O) \in \Gamma \times SO(2)\},$$
 where the closure is with respect to the norm $\|\cdot\|_{\ltwo}$ on $Q$. The shape space consequently is defined as $\mathcal{Q}_s:=\{[Q_\C]: Q_\C \in \mathcal{Q}\}$. Property (2) ensures that the metric induced by the norm $\|\cdot\|_{\ltwo}$ on $\mathcal{Q}$ descends onto a metric $d$ on the shape space (quotient space) $\mathcal{Q}_s$ in a natural way. Given two curves $\C_1$ and $\C_2$, the shape distance between them is defined as
 \begin{align}
 \label{distance}
 d(\C_1,\C_2)&:=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|Q_{\C_1} - OQ_{\C_2}(\gamma)\sqrt{\gamma'}\|_{\ltwo}\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)}\Big[\int_{\uc}\left\|Q_{C_1}(t)-OQ_{C_2}(\gamma(t))\sqrt{\gamma'(t)}\right\|_2^2dt\Big]^{1/2} \nonumber\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|OQ_{C_1}(\gamma)\sqrt{\gamma'} - Q_{C_2}\|_{\ltwo}\nonumber.
  \end{align}
  The symmetry with respect to the action either on $Q_{C_1}$ or $Q_{C_2}$ is an attractive feature and will be used profitably in the sequel.

\section{Preserving scale}
For an absolutely continuous curve $\C:\uc \to \Rtwo$ consider the transformation 
$$\C \mapsto C'=:Q^{\star}_\C = Q_\C\|C'\|_{\ltwo} $$ 
where $\C'$ is the (vector) derivative of $\C(t)$ with respect to $t$.

In this case, the unique (again, up to translations) inverse of this curve $Q^{\star}_C$ is $\int_0^t Q^{\star}_C(s)ds$. To ensure that the curves are closed we need to impose an additional constraint: $\int_\uc Q^{\star}_C(t)=0$. 
  Thus the image of the set of absolutely continuous closed curves , $ \{\C:\uc \to \Rtwo: \int_\uc \|C'(t)\|_2dt=l\}$, under the scale preserving transform map $\C \mapsto Q^{\star}_\C$ is the set
\[
\mathcal{Q}:=\Big\{Q^{\star}_\C:  \int_\uc Q^{\star}_C(t)dt=0\Big\},
\]

 Starting with a curve $\C$ we can now define its size and shape to be the equivalence class or its orbit of its corresponding velocity transform:
 $$[Q^{\star}_\C]=\text{closure}\{OQ^{\star}_\C(\gamma) \sqrt{\gamma'}: (\gamma,O) \in \Gamma \times SO(2)\},$$
 where the closure is with respect to the norm $\|\cdot\|_{\ltwo}$ on $Q^{\star}$.

 In the square-root velocity framework, dividing the velocity function by its norm gives us a function that is of norm one.  This implies that it is a point on the unit sphere.  In the finite-dimensional vecotr set-up this amounts to keeping just the direction of the vectors and discarding all other information and the distance between two curves can then be viewed as the angle between these two vectors.  However, when preserving scale this defintion of distance will need to be modified as both the angle AND magnitude of the vectors are both of interest.

 Therefore, a modified defintion of distance that preserves scaling between two curves will be defined:

 %Is this one symmetric?  
 \begin{align}
 \label{distance}
 d^{\star}(\C_1,\C_2)&:=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|Q^{\star}_{\C_1} - OQ^{\star}_{\C_2}(\gamma)\sqrt{\gamma'}\|_{\ltwo}\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)}\Big[\int_{\uc}\left\|Q^{\star}_{C_1}(t)-OQ^{\star}_{C_2}(\gamma(t))\sqrt{\gamma'(t)}\right\|_2^2dt\Big]^{1/2} \nonumber\\
 &=\inf_{(\gamma,O) \in \Gamma \times SO(2)} \|OQ^{\star}_{C_1}(\gamma)\sqrt{\gamma'} - Q^{\star}_{C_2}\|_{\ltwo}\nonumber.
  \end{align}

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Missingness Function}
Now consider a function, $R$, defined over the set $\mathbb{S}$, where $R(t) = 1$ if the shape function is observed at $t \in \mathbb{S}$ and $R(t) = 0$ otherwise (i.e. the shape function is observaed at $t \in \mathbb{S}$)

Specifically,   
$$
R: \mathbb{S} \rightarrow \left\{0,1\right\}^2
$$

In $\Rtwo$, for a given point $t$ we have $C(t) = (x_t, y_t)$ and the defintion of the function $R$ allows these two be missing individually (e.g. $x_t$ could be observed and $y_t$ is missing or $y_t$ could be observed and $x_t$ is missing), in our setting both the $x_t$ and $y_t$ values are either both observed or both missing and so specifically in this case the function is of the following form: 
$$
R: \mathbb{S} \rightarrow \left\{\{0,0\},\{1,1\}\right\}
$$

Now for a specific curve $C$ in the equivalence class $[\C]$ (or $[\C^{\star}]$), we can define the function $R$ over $\mathbb{S}$.  However, for every reparameterization of $C$ in the equivalence class $[\C]$ (or $[\C^{\star}]$), there is a corresponding reparameterization of $R$.  Also, note that $R^{-1}(1)$ is the subset of domain $\mathbb{S}$ such that the coordinates of the shape are observed for every point in this set.  Further, here only the case where $R^{-1}(1)$ be a connected subset of $\mathbb{S}$ is considered, though in general, this does not have to be the case. This prevents us from considering partially observed shapes with multiple ``gaps" (though this is an interesting idea for future work).

For a given shape, there are many possible reparameterizations of the missingness function, and an equivalence class can be defined for the missingness function across reparametertizations as follows:

%I think I only need reparameterizations for R?  Rotations too?  
$$
[\R] = \{ R(\gamma(t)) , \gamma \in \Gamma \} 
$$.

Note that for every curve $\C$ and corresponding $R$, in order to maintain the correct relationship, if $\C$ is reparameterized by $\gamma \in \Gamma$
 $[\C]$ there is a correspondo

\section{Curve completion and classification}
Suppose we are given closed planar curves $\C^m_j:\uc \to \mathbb{R}^2,j=1\ldots,n$ each of which has been observed only on a region $\R_j \subset \uc$, where $\uc$ is the unit circle in $\Rtwo$. We assume that the curves are absolutely continuous. Additionally, a training sample $\{(y_i,\C^o_i),1,\ldots,N\}$ consisting of class labels $y_i \in \{0,1\}$ and $C^o_i:\uc\to \mathbb{R}^2$, fully observed on a common domain $\uc$, is provided. The set of fully observed curves are elements of the set $\mathcal{C}$; denote by $\mathcal{C}^m$ the set of all partially observed curves.
